---
title: "Correlation Analysis"
author: "Felipe Barros"
date: "12/7/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Correlation Analysis

Tasks:  

> Ou seja, entre riqueza e abund창ncia do msm grupo, entre riqueza de grupos diferentes, entre abund창ncia de grupos diferentes e tb entre riqueza e abund창ncia de grupos diferentes. Os pontos s찾o para resgatar os valores preditos dos modelos.

```{r loading_packages, warning=FALSE, include=FALSE}
library(raster)
library(readr)
library(dplyr)
library(tmap)
library(sf)
library(corrplot)
#install.packages("ape") 
library(ape)
#install.packages("pgirmess")
library(pgirmess)
```
```{r loading_results, warning=FALSE, echo=FALSE, results='hide'}
# Getting all results
indexLayers <- list.files("../Data/Raster/Results", pattern = "Abundance_Norm.tif$|ness_Norm.tif$", recursive = TRUE, full.names = TRUE)

# stacking
stackedIndex <- stack(indexLayers)
#mask <- stackedIndex[[1]]>0

cor.mtest <- function(mat, ...) {
    mat <- as.matrix(mat)
    n <- ncol(mat)
    p.mat<- matrix(NA, n, n)
    diag(p.mat) <- 0
    for (i in 1:(n - 1)) {
        for (j in (i + 1):n) {
            tmp <- cor.test(mat[, i], mat[, j], ...)
            p.mat[i, j] <- p.mat[j, i] <- tmp$p.value
        }
    }
  colnames(p.mat) <- rownames(p.mat) <- colnames(mat)
  p.mat
}

```

## Sampling pixels over all models  

```{r sampling_overall, eval = FALSE, echo=TRUE}
# sampling over all layers
set.seed(031219)
samp <- sampleRandom(stackedIndex, 50000, xy = TRUE, sp=FALSE, na.rm = TRUE)
samp <- as_tibble(samp)
write_csv(samp, "../Data/CSV/Results/Rastersamples.csv")
```

```{r loading_sampleOverall, message=FALSE}
samp <- read_csv("../Data/CSV/Results/Rastersamples.csv")
samples_sf <- st_as_sf(samp, coords = c("x", "y"), crs = 4326)

#tm_shape(mask) +
#  tm_raster( legend.show = FALSE) +
#  tm_shape(samples_sf) + 
#  tm_dots()

```  

### Correlation over all models  

```{r correlationOverall, message=FALSE}
sampCor <- cor(samp[,3:ncol(samp)], method="pearson")
corrplot(sampCor, type="upper", order="hclust", col=RColorBrewer::brewer.pal(n=8, name="RdYlBu"), cl.align.text = "l", addCoef.col = "black", sig.level = 0.05, diag = F, insig = "blank", p.mat = sampCor)
```  

```{r}
# matrix of the p-value of the correlation
p.samp <- cor.mtest(samp[,-c(1,2)])
pander::pander(p.samp, caption = "p-value for each pair of variable", split.tables	= Inf, style = 'rmarkdown')
```


#### Spatial correlation:  

```{r}
#https://stats.idre.ucla.edu/r/faq/how-can-i-calculate-morans-i-in-r/
#https://gis.stackexchange.com/questions/304122/compute-morans-i-on-more-than-one-attribute

```

```{r spatGeneralCorrelation, message=FALSE, error=FALSE, warning=FALSE, echo=FALSE, prompt=FALSE, comment=FALSE}
samp.dists <- as.matrix(dist(cbind(samp$x, samp$y)))
samp.dists <- 1/samp.dists
diag(samp.dists) <- 0
#pander::pander(samp.dists[1:5, 1:5], caption = "First 5 values of distances in degree")
```  

**I-Moran** calculated for:  

* Flora_Abundance_Norm = `r Moran.I(samp$Flora_Abundance_Norm, samp.dists)$observed`  
  * p-value = `r Moran.I(samp$Flora_Abundance_Norm, samp.dists)$p.value`  
* Flora_Richness_Norm = `r Moran.I(samp$Flora_Richness_Norm, samp.dists)$observed`  
  * Flora_Richness_Norm = `r Moran.I(samp$Flora_Richness_Norm, samp.dists)$p.value`   
* Vert_Abundance_Norm = `r Moran.I(samp$Vert_Abundance_Norm, samp.dists)$observed`   
  * Vert_Abundance_Norm = `r Moran.I(samp$Vert_Abundance_Norm, samp.dists)$p.value`  
* Vert_Richness_Norm = `r Moran.I(samp$Vert_Richness_Norm, samp.dists)$observed`  
  * Vert_Richness_Norm = `r Moran.I(samp$Vert_Richness_Norm, samp.dists)$p.value`  
* Invert_Abundance_Norm = `r Moran.I(samp$Invert_Abundance_Norm, samp.dists)$observed`  
  * Invert_Abundance_Norm = `r Moran.I(samp$Invert_Abundance_Norm, samp.dists)$p.value`  
* Invert_Richness_Norm = `r Moran.I(samp$Invert_Richness_Norm, samp.dists)$observed`  
  * Invert_Richness_Norm = `r Moran.I(samp$Invert_Richness_Norm, samp.dists)$p.value`  

**Correlogram**
```{r spatGeneralCorrelogram, message=FALSE, error=FALSE, warning=FALSE, echo=TRUE, prompt=FALSE, comment=FALSE}

#t <- correlog.nc(x = samp$x, y = samp$y, z = samp[, 5], w = NULL, increment = 100, resamp = 1000, latlon = TRUE, na.rm = FALSE, quiet = FALSE)
t <- pgirmess::correlog(coords=samp[,c(1:2)], z=samp$Flora_Abundance_Norm, method="Moran", nbclass=NULL)
plot(t)
```  

```{r, eval = FALSE}
#install.packages("devtools")
if (!require("rspatial")) devtools::install_github('rspatial/rspatial')
library(rspatial)
ppp(Ameacadas_SEM_14_12[,3], Ameacadas_SEM_14_12[,4], marks=as.factor(Ameacadas_SEM_14_12[,2]), window=as.owin(area_sem))
```  

## Richness between groups  

> entre riqueza de grupos diferentes
 
```{r richnessBetweenGroups, eval=FALSE}
set.seed(031218)
sampRichnessBG <- sampleRandom(stackedIndex[[c(2,4,6)]], 50000, xy = TRUE, sp=FALSE, na.rm = TRUE)
sampRichnessBG <- as_tibble(sampRichnessBG)
write_csv(sampRichnessBG, "../Data/CSV/Results/RichnessRastersamples.csv")
```
```{r correlationRichness, message=FALSE}
sampRichnessBG <- read_csv("../Data/CSV/Results/RichnessRastersamples.csv")
sampCorRichness <- cor(sampRichnessBG[,3:ncol(sampRichnessBG)], method="pearson")
corrplot(sampCorRichness, method="circle", type="upper", diag=F, order = "hclust", col=RColorBrewer::brewer.pal(n=8, name="RdYlBu"), cl.align.text = "l", addCoef.col = "black")
```  

```{r}
# matrix of the p-value of the correlation
p.sampRichnessBG <- cor.mtest(sampRichnessBG[,-c(1,2)])
pander::pander(p.sampRichnessBG, caption = "p-value for each pair of variable", split.tables	= Inf, style = 'rmarkdown')
```

## Abundance between groups  

> abundancia de grupos diferentes
 
```{r abundanceBetweenGroups, eval=FALSE}
# Flora 
set.seed(031217)
sampAbundanceBG <- sampleRandom(stackedIndex[[c(1,3,5)]], 50000, xy = TRUE, sp=FALSE, na.rm = TRUE)
sampAbundanceBG <- as_tibble(sampAbundanceBG)
write_csv(sampAbundanceBG, "../Data/CSV/Results/AbundanceRastersamples.csv")
```
```{r correlationAbundance, message=FALSE}
sampAbundanceBG <- read_csv("../Data/CSV/Results/AbundanceRastersamples.csv")
sampCorAbundance <- cor(sampAbundanceBG[,3:ncol(sampAbundanceBG)], method="pearson")
corrplot(sampCorAbundance, method="circle", type="upper", diag=F, order = "hclust", col=RColorBrewer::brewer.pal(n=8, name="RdYlBu"), cl.align.text = "l", addCoef.col = "black")
```  

```{r}
# matrix of the p-value of the correlation
p.sampAbundanceBG <- cor.mtest(sampAbundanceBG[,-c(1,2)])
pander::pander(p.sampAbundanceBG, caption = "p-value for each pair of variable", split.tables	= Inf, style = 'rmarkdown')
```

## Correlation by group

### Flora  

```{r richnessAbudanceFlora, eval=FALSE}
# Flora 
set.seed(031216)
sampFlora <- sampleRandom(stackedIndex[[c(1:2)]], 50000, xy = TRUE, sp=FALSE, na.rm = TRUE)
sampFlora <- as_tibble(sampFlora)
write_csv(sampFlora, "../Data/CSV/Results/FloraRastersamples.csv")
```
```{r correlationFlora, message=FALSE}
sampFlora <- read_csv("../Data/CSV/Results/FloraRastersamples.csv")
sampCorFlora <- cor(sampFlora[,3:ncol(sampFlora)], method="pearson")
corrplot(sampCorFlora, method="circle", type="upper", diag=F, order = "hclust", col=RColorBrewer::brewer.pal(n=8, name="RdYlBu"), cl.align.text = "l", addCoef.col = "black")
```  

P-value from `cor.test()` = `r cor.test(sampFlora$Flora_Abundance_Norm, sampFlora$Flora_Richness_Norm)$p.value`, correlation value = `r cor.test(sampFlora$Flora_Abundance_Norm, sampFlora$Flora_Richness_Norm)$estimate`.  

### Invertebrates  

```{r richnessAbudanceInvert, eval=FALSE}
# Vert
set.seed(031215)
sampInvert <- sampleRandom(stackedIndex[[c(3:4)]], 50000, xy = TRUE, sp=FALSE, na.rm = TRUE)
sampInvert <- as_tibble(sampInvert)
write_csv(sampInvert, "../Data/CSV/Results/InvertRastersamples.csv")
```
```{r correlationInvert, message=FALSE}
sampInvert <- read_csv("../Data/CSV/Results/InvertRastersamples.csv")
sampCorInvert <- cor(sampInvert[,3:ncol(sampInvert)], method="pearson")
corrplot(sampCorInvert, method="circle", type="upper", diag=F, order = "hclust", col=RColorBrewer::brewer.pal(n=8, name="RdYlBu"), cl.align.text = "l", addCoef.col = "black")
```  

P-value from `cor.test()` = `r cor.test(sampInvert$Invert_Abundance_Norm, sampInvert$Invert_Richness_Norm)$p.value`, correlation value = `r cor.test(sampInvert$Invert_Abundance_Norm, sampInvert$Invert_Richness_Norm)$estimate`.   

### Vertebrates  

```{r richnessAbudanceVert, eval=FALSE}
# Vert
set.seed(031214)
sampVert <- sampleRandom(stackedIndex[[c(5:6)]], 50000, xy = TRUE, sp=FALSE, na.rm = TRUE)
sampVert <- as_tibble(sampVert)
write_csv(sampVert, "../Data/CSV/Results/VertRastersamples.csv")
```
```{r correlationVert, message=FALSE}
sampVert <- read_csv("../Data/CSV/Results/VertRastersamples.csv")
sampCorVert <- cor(sampVert[,3:ncol(sampVert)], method="pearson")
corrplot(sampCorVert, method="circle", type="upper", diag=F, order = "hclust", col=RColorBrewer::brewer.pal(n=8, name="RdYlBu"), cl.align.text = "l", addCoef.col = "black")
```  

P-value from `cor.test()` = `r cor.test(sampVert$Vert_Abundance_Norm, sampVert$Vert_Richness_Norm)$p.value`, correlation value = `r cor.test(sampVert$Vert_Abundance_Norm, sampVert$Vert_Richness_Norm)$estimate`.   
